---
title: Extracting Data from Rasters
date: 2015-01-10T07:58:18+00:00
guid: http://rampages.us/dyerlab/?p=278
permalink: /2015/01/10/extracting-data-from-rasters/
categories:
  - Code
  - Tutorials
tags:
  - Araptus attenuata
  - Baja California
  - ggplot2
  - R
  - raster
---
This document shows you how to extract data from rasters.

&nbsp;

## Getting The Libraries

First, I'll load in some packages to get the ability to work with raster data and to load in the _Arapatus attenuatus_ data set (it is part of the default **gstudio** package).

```{r}
require(raster)
require(gstudio)
```


## Loading and Cropping Rasters

We can load in the raster, and then crop it to just the are we need. These rasters were downloaded from [[http://www.worldclim.org](http://www.worldclim.org/)] and are much larger than the study area. This just makes it easier on the computer to not have to deal with such large areas. After cropping it, we will load in the annual precip and temperature data as well.

```{r}
tmp <- raster("alt_22.tif")
e <- extent(c(-115, -109, 22, 30))
baja_alt <- crop(tmp, e)
baja_temp <- crop(raster("bio1_22.tif"), e)
baja_prec <- crop(raster("bio12_22.tif"), e)
```


## Getting Example Data from _Araptus attenuatus_

Now, lets grab the _Araptus_ data and look at the data and plot out the locations.

```{r}
library(ggplot2)
data(arapat)
summary(arapat)
coords <- strata_coordinates(arapat)
```



```{r eval = FALSE}
##
## This no longer works UNLESS you have a google API key
## 
library(ggmap)
map <- population_map(coords, map.source = "stamen")
ggmap( map ) + geom_point(aes(x=Longitude,y=Latitude), data=coords)
```


## Extracting Point Data

To elevation, temperature and precipitation from the rasters for each sampling location, we need to translate them into points first. I'll first grab the coordinate data as a _data.frame_.


Then we can grab them using the normal functions in the **sp** library.

```{r}
pts <- SpatialPoints(coords[, 2:3])
coords$elev <- extract(baja_alt, pts)
coords$temp <- extract(baja_temp, pts)
coords$prec <- extract(baja_prec, pts)
coords
```


## Plotting Trend lines.

Cool, lets sort this by latitude

```{r}
coords <- coords[order(coords$Latitude), ]
```

and then plot out some values to look at what is going on.

```{r}
ggplot(coords, aes(x = Latitude, y = elev)) + geom_line(color = "gray") + theme_bw() + geom_text(aes(y = elev + 10, label = Stratum), color = "blue")
```


```{r}
ggplot(coords, aes(x = Latitude, y = temp)) + geom_line(color = "gray") + theme_bw() + geom_text(aes(y = temp + 5, label = Stratum), color = "blue")
```


```{r}
ggplot(coords, aes(x = Latitude, y = prec)) + geom_line(color = "gray") + theme_bw() + geom_text(aes(y = prec + 10, label = Stratum), color = "blue")
```

